---
title: "2 - Content description of the Phenolics MS/MS repository"
author: 
  - name: Cristian Quiroz-Moreno, Jessica Coopestone
    affiliation:
    - Horticulture and Crop Science Department, Ohio State University, US 
    email: quirozmoreno.1@osu.edu
output: 
  BiocStyle::html_document:
    self_contained: yes
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: show
date: "`r doc_date()`"
package: "`r pkg_ver('PhenolicsDB')`"
vignette: >
  %\VignetteIndexEntry{2 - Content description of the Phenolics MS/MS repository}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    crop = NULL ## Related to https://stat.ethz.ch/pipermail/bioc-devel/2020-April/016656.html
)
```


```{r vignetteSetup, echo=FALSE, message=FALSE, warning = FALSE}
## Track time spent on making the vignette
startTime <- Sys.time()
library('dplyr')
library('Rdisop')
library('ggplot2')
## Bib setup
library("RefManageR")

## Write bibliography information
bib <- c(
    R = citation(),
    BiocStyle = citation("BiocStyle")[1],
    knitr = citation("knitr")[1],
    RefManageR = citation("RefManageR")[1],
    rmarkdown = citation("rmarkdown")[1],
    sessioninfo = citation("sessioninfo")[1],
    testthat = citation("testthat")[1],
    PhenolicsDB = citation("PhenolicsDB")[1]
)

```

# Package content description


This package is intended to archive the raw mzml files of authentic
standards, as well as structured data frame aimed to be used jointly
with [MS2extract](https://cooperstonelab.github.io/MS2extract/), to
automate the process of creating in-house MS/MS libraries.

This database contains majorly phenolics compounds, aimed to decipher
and annotate the *Rosacea* metabolome. However, as phenolics are
expressed in a wide range of organisms, this database can be used in
multiples scenarios.

We grouped the total database in two batches. The batches are based on
the instrument we used to acquire the MS/MS data. For the first batch,
we used an Agilent QTOF *6545*, and for the second batch, we used an
Agilent QTOF *6546*. In both batches, we collected targeted MS/MS using
the chromatographic method reported in [Bilbrey at al.
(2022)](https://nph.onlinelibrary.wiley.com/doi/full/10.1111/nph.17693).
We collected data in negative and positive polarity, and we used 20 and
40 eV (CE) using collision induced dissociation (CID). Additionally, in
the second batch, we also collected data at 60 and 80 eV (CE). The
following table describe what you expect to find in this database.

### First Batch (Agilent QTOF 6545)

| Polarity | Collision Energy (eV) | read_dt table               | metadata table           |
|--------------|-------------|---------------------|----------------------|
| Positive | 20                    | \`data("read_pos20_6545")\` | data("metdt_pos20_6545") |
| Positive | 40                    | \`data("read_pos40_6545")\` | data("metdt_pos40_6545") |
| Negative | 20                    | \`data("read_neg20_6545")\` | data("metdt_neg20_6545") |
| Negative | 40                    | \`data("read_neg40_6545")\` | data("metdt_neg40_6545") |

### Second Batch (Agilent QTOF 6546)

| Polarity | Collision Energy (eV) | read_dt table                 | metadata table           |
|--------------|-------------|----------------------|----------------------|
| Positive | 20                    | \`data("read_pos20_6546")\`   | data("metdt_pos20_6546") |
| Positive | 40                    | \`data("read_pos40_6546")\`   | data("metdt_pos40_6546") |
| Positive | 60                    | \`data("read_pos6080_6546")\` | data("metdt_pos60_6546") |
| Positive | 80                    | \`data("read_pos6080_6546")\` | data("metdt_pos80_6546") |
| Negative | 20                    | \`data("read_neg20_6546")\`   | data("metdt_neg20_6546") |
| Negative | 40                    | \`data("read_neg40_6546")\`   | data("metdt_neg40_6546") |
| Negative | 60                    | \`data("read_neg60_6546")\`   | data("metdt_neg60_6546") |
| Negative | 80                    | \`data("read_neg80_6546")\`   | data("metdt_neg80_6546") |

## List of metabolites

In the `extdata` directory you will find all the `.mzml` files for all
standards, as well as the structure tables to be used jointly with
[MS2extract](https://cooperstonelab.github.io/MS2extract/).

In the following table you can see the list of metabolites we are
including in this version of the database, as well as in what batch you
will find them.

> Disclaimer: Metabolites with no reported data file have not MS/MS data

```{r echo=FALSE, message=FALSE, warning=FALSE}

data("read_neg20_6545", package = "PhenolicsDB")
batchRead_Neg20 <-   dplyr::mutate(read_neg20_6545, Batch = "(1) QTOF-6545")

data("read_neg20_6546", package = "PhenolicsDB")
read_neg20_6546 <-  dplyr::mutate(read_neg20_6546, Batch = "(2) QTOF-6546")

all_std <- dplyr::bind_rows(batchRead_Neg20, read_neg20_6546)  


all_std <- dplyr::select(all_std, -File, -Ionization_mode)
all_std
```

# Phytochemicals in two dimensions

As it was explained above, the chromatographic method employed for the 
LC separation is described in [Bilbrey at al.
(2022)](https://nph.onlinelibrary.wiley.com/doi/full/10.1111/nph.17693).

Therefore, we can display where each phytochemical elutes in terms of
retention time and *m/z*.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
data("smiles2classyfire", package = "PhenolicsDB")

title_text <- "Distribution of the metabolites included in PhenolicsDB"
subtitle_text <- paste("Metabolites are colored based on", 
                        "Superclass information from Classyfire")

all_std <- dplyr::mutate(all_std, 
              Monoisotopic = sapply(all_std$Formula,
                                    FUN = function(x){
                                      Rdisop::getMolecule(x)$exactmass
                                    } ) )  %>%  dplyr::rowwise() %>% 
  dplyr::mutate(rt = mean(c(min_rt, max_rt) )/60) %>% 
  left_join(smiles2classyfire)


std_distribution <- all_std %>%
  ggplot2::ggplot( ggplot2::aes(rt, Monoisotopic, 
                                color = Superclass_CF,
                                shape = Superclass_CF)) +
  ggplot2::geom_point(alpha = 0.8, size = 2.5) + 
  ggplot2::theme_light() + 
  ggsci::scale_color_frontiers()  +
  ggplot2::theme(legend.position = c(0.82, 0.77), 
                legend.background = element_rect(fill = "white",
                                                  color = "black")) +
  ggplot2::labs(x = "Retention time (min)",
                y = "Neutral monoisotopic mass (Da)",
                title = title_text,
                subtitle = subtitle_text)

std_distribution
```


# Information about this vignette

Code for creating the vignette

```{r createVignette, eval=FALSE}
## Create the vignette
library("rmarkdown")
system.time(render("Content.Rmd", "BiocStyle::html_document"))

## Extract the R code
library("knitr")
knit("Content.Rmd", tangle = TRUE)
```

Date the vignette was generated.

```{r reproduce1, echo=FALSE}
## Date the vignette was generated
Sys.time()
```

Wallclock time spent generating the vignette.

```{r reproduce2, echo=FALSE}
## Processing time in seconds
totalTime <- diff(c(startTime, Sys.time()))
round(totalTime, digits = 3)
```

`R` session information.

```{r reproduce3, echo=FALSE}
## Session info
library("sessioninfo")
options(width = 120)
session_info()
```



# Bibliography

This vignette was generated using `r Biocpkg("BiocStyle")` `r Citep(bib[["BiocStyle"]])`
with `r CRANpkg("knitr")` `r Citep(bib[["knitr"]])` and `r CRANpkg("rmarkdown")` `r Citep(bib[["rmarkdown"]])` running behind the scenes.

Citations made with `r CRANpkg("RefManageR")` `r Citep(bib[["RefManageR"]])`.

```{r vignetteBiblio, results = "asis", echo = FALSE, warning = FALSE, message = FALSE}
## Print bibliography
PrintBibliography(bib, .opts = list(hyperlink = "to.doc", style = "html"))
```
